---
title: "SQL in R"
author: "Jessica Jagdeo"
date: "3/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      connection = lter_db)

library(tidyverse)
library(here)
library(DBI)
library(RSQLite)
```

### 1. Connect to the database and tell R that we're using SQLite:

```{r}

lter_db <- DBI::dbConnect(drv = RSQLite::SQLite(),
                          dbname = here("data", "marine.sqlite")) 

# Tells R what engine you're going to use when doing SQL querying (SQLite) and tells us what database to connect to (marine.sqlite)

```

### 2. What tables exist in the database?

To check which tables exist (table names) in a database, use 'DBI::dbListTables()':

```{r}

DBI::dbListTables(lter_db)

# Tables are fish_data, inverts_data, and lobsters_data

```

### 3. Which variables are in a database table?

Use 'DBI::dbListFields()':

```{r}

DBI::dbListFields(lter_db, "fish_data")

# Column names for fish_data are year, site, common_name, and total_count

```

```{r}

DBI::dbListFields(lter_db, "inverts_data")

# Column names are month, site, common_name, 2016, 2017, 2018

```

#### Read in a database table as an R data frame

Use 'DBI::dbReadTable()':

```{r}

fish_df <- DBI::dbReadTable(lter_db, "fish_data")
class(fish_df) # Check to see if it's a data frame

```

#### Query a database using SQL

- SELECT: Gets columns
- FROM: What table in a database to get something from
- Use * to indicate "get all"

```{sql connection=lter_db}

SELECT * FROM fish_data

```

To get a subset of the columns, we can just list them after SELECT, separated by a comma:

```{sql connection=lter_db}

SELECT year, site FROM fish_data

```

We might want to only query for ROWS (observations) that match a given condition we set.

- In R: dplyr::filter()
- In SQL: WHERE

```{sql connection=lter_db}

SELECT * FROM fish_data WHERE year == 2016

```

Now, let's query to keep all columns from fish_data, but only if the common_name variable matches "garibaldi"

```{sql connection=lter_db}

SELECT common_name, site FROM fish_data WHERE common_name == "garibaldi"

```

In dplyr: 
filter for an "and" statement: , &

In SQL: "and"

Example: Query to only keep observations where the common_name == "garibaldi" AND site == "mohk"

```{sql connection=lter_db}

SELECT * FROM fish_data WHERE common_name == "garibaldi" AND site == "mohk"

```

In R, to write an OR statement:
- |
- %in%

In SQL: Use OR

Example: Query to keep observations where year is 2017 OR site is "abur"

```{sql connection=lter_db}

SELECT * FROM fish_data WHERE year == 2017 OR site == "abur"

```

#### Store an output from SQL querying

Add in the code chunk header `output.var = "name_object"`

```{sql connection=lter_db, output.var = "fish_2017_df"}

SELECT * FROM fish_data WHERE year == 2017

```

Query using SQL from an R code chunk

`DBI::dbSendQuery()`

```{r}

garibaldi <- dbSendQuery(lter_db, "SELECT * FROM fish_data WHERE common_name == 'garibaldi'")

gar_df <- dbFetch(garibaldi)

```

### Querying using dplyr-ish code

- `dbplyr`

```{r}

fish_tbl <- tbl(lter_db, "fish_data")

# Some wrangling:

yearly_fish <- fish_tbl %>% 
  group_by(year) %>% 
  summarize(
    total_fish = sum(total_count)
  )

yearly_fish %>%  show_query() # Shows you the SQL code necessary to replicate what you did above in R dplyr

yearly_fish %>%  collect() # Get the results of the SQL query

```

#### Adding or removing tables from a .sqlite database

Use `DBI::dbWriteTable` and `DBI::dbRemoveTable` functions to add or remove tables from/to an existing database

```{r}

dbWriteTable(lter_db, "mtcars_df", mtcars)

# Check to see what tables are in the database now:

DBI::dbListTables(lter_db)

# Remove a table using dbRemoveTables

DBI::dbRemoveTable(lter_db, "mtcars_df")

```

